---
title: "Lab 1 - Data visualization"
author: "Saisha Dhar"
format: pdf
editor: visual
---

## Load Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(viridis)

```

## Exercise 1

```{r}

#| label: histogram-population-density-of-counties

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 10000) + labs(x = "Population Density", 
      y = "Counties") + ggtitle("Population Density of Counties")

```
The shape of the distribution is skewed right, which means there are more counties that have lower rather than higher population density. There appear to be a couple outliers, as there are some counties that are between 75,000-100,000 and they aren't in the highest frequency categories.


## Exercise 2

```{r}

#| label: degree-versus-poverty-scatterplot

ggplot(midwest, aes(x = percollege, y = percbelowpoverty, color = state)) +
  geom_point() + 
  scale_fill_viridis_d() +
  labs(x = "Percentage of People with a College Degree", 
       y = "Percentage of People Below Poverty", color = "State") + 
  ggtitle("Percentage of People with a College Degree vs. 
          Percentage of People Below Poverty")

```

## Exercise 3

I observe a relatively neutral general distribution, I don't see a very strong negative or positive association between the percentage of people with a college degree compared to the percentage of people below the poverty line. When looking at states, however, some have more of a negative association like WI and IL, and states like IN and MI are more of a neutral association.  

## Exercise 4

```{r}
#| label: generate-scatterplots-by-state

ggplot(midwest, aes(x = percollege, y = percbelowpoverty, color = state)) +
  geom_point() + 
  scale_fill_viridis_d() + 
  geom_smooth(se = FALSE) + 
  facet_wrap(~state) +  
  labs(x = "% of People with a College Degree", 
       y = "% of People Below Poverty", color = "State") + 
  ggtitle("% of People with a College Degree 
          vs. % of People Below Poverty by State")

```

I prefer this plot as compared to exercise 2 because I am able to see the relationship of percent of people with a college degree vs. percent of people below the poverty line more clearly in different states, while exercise 2 I can't see the relationships that clearly because of too many points together. 


## Exercise 5

```{r}

#| label: boxplot-area-of-counties-by-state

ggplot(midwest, aes(x = state, y = area, fill = state)) + 
  geom_boxplot() +
labs(x = "State", y = "Area of County", 
     fill = "State") + 
  ggtitle("Area of Counties by State")

```
Each boxplot represents the areas of the counties within a state. Indiana and Ohio have counties with a smaller and lower range of areas, while the other states have larger quartiles, ranges, and areas overall. Michigan is the state with the single largest county, because each point is the area of a county and MI has the point highest on the y-axis. 

## Exercise 6


```{r}
#| label: generate-segmented-bar-chart

midwest <- midwest |>
  mutate(metro = if_else(inmetro == 1, "Yes", "No"))
ggplot(data = midwest, aes(x = state, fill = metro, y=1)) +
  geom_bar(position = "fill", stat = "identity") +
labs(x = "State", y = "Percentage of Counties in Metro", 
     fill = "In Metro") +
ggtitle("Counties in Metro Area by State")

```

## Exercise 7

```{r}
#| label: generate-college-degree-plot

ggplot(midwest, aes(x = percollege, y=popdensity, color=percbelowpoverty)) +
  geom_point(alpha = 0.5, size = 2) + facet_wrap(~state) + 
  labs(x = "% college educated", y = "Population density (person/unit area)", 
       color = "% below \n poverty line") + 
  ggtitle("Do people with college degrees tend to live in 
          denser areas?") + 
  theme_minimal()

```

